---
title: "Raw count of crime by type across selected Victorian suburbs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 1600)
```

The crime data files are downloaded from the [Victoria Government Crime Statistics website](https://www.crimestatistics.vic.gov.au/crime-statistics/latest-crime-data/recorded-criminal-incidents-0). Go to the bottom of the page and select "Criminal incidents by principal offence and postcode or suburb/town". Unfortunately, the data is behind a Tableau plugin in the browser and cannot be deep-linked, so I had to manually select the suburbs I wanted in the dropdown, sort the data by a column (any column will do; otherwise the data cannot be downloaded in CSV), and click download at the bottom of the page. Click "Data", and be sure to click on the "Full data" tab.

Now, the analysis is shown below.

## Steps
Read all suburb files and combine them into one big data frame
```{r}
df <- data.frame()
files <- c("CAULFIELD", "CHADSTONE", "BRUNSWICK", "NORTHCOTE", "COBURG", "SEDDON", "SUNSHINE", "HADFIELD", "PASCOE_VALE", "RESERVOIR", "CLAYTON", "BOX_HILL", "SUNBURY", "PRESTON", "FAWKNER", "BROADMEADOWS", "DANDENONG")
for (f in files) {
  df <- rbind(df, read.csv(paste("./crime/", f, ".csv", sep = ""), sep = "\t"))
}
```
Total number of rows:
```{r}
nrow(df)
```
Columns in the data frame:
```{r}
names(df)
```

The offence divisions are:
```{r}
offence_divisions <- c()
division <- unique(df[,"Offence.Division"])
num_of_divisions <- length(division)
for (i in 1:num_of_divisions) {
  offence_divisions <- c(offence_divisions, levels(division)[i])
}
offence_divisions <- sort(offence_divisions)
offence_divisions
```

The offence subdivisions are:
```{r}
offence_subdivisions <- c()
subdivision <- unique(df[,"Offence.Subdivision"])
num_of_subdivisions <- length(subdivision)
for (i in 1:num_of_subdivisions) {
  offence_subdivisions <- c(offence_subdivisions, levels(subdivision)[i])
}
offence_subdivisions <- sort(offence_subdivisions)
offence_subdivisions
```

The suburbs are:
```{r}
suburbs <- c()
suburb_names <- unique(df[,"Suburb.Town.Name"])
num_of_suburbs <- length(suburb_names)
for (i in 1:num_of_suburbs) {
  suburbs <- c(suburbs, levels(suburb_names)[i])
}
suburbs <- sort(suburbs)
suburbs
```

```{r}
zero_missing_suburb_division <- function (df, year) {
  df$Offence.Subdivision <- NULL
  df$Postcode <- NULL
  for (suburb in suburbs) {
    for (div in offence_divisions) {
      df <- rbind(df, data.frame(Incidents.Recorded = 0, Offence.Division = div, Suburb.Town.Name = suburb, Year.ending.December = year))
    }
  }
  df
}

zero_missing_suburb_subdivision <- function (df, year) {
  df$Offence.Division <- NULL
  df$Postcode <- NULL
  for (suburb in suburbs) {
    for (subdiv in offence_subdivisions) {
      df <- rbind(df, data.frame(Incidents.Recorded = 0, Offence.Subdivision = subdiv, Suburb.Town.Name = suburb, Year.ending.December = year))
    }
  }
  df
}

```

Counts of crime division by suburb for each year:
```{r}
for (year in rev(sort(unique(df$Year.ending.December)))) {
  df_by_year <- df[df$Year.ending.December == year, ]
  df_by_year <- zero_missing_suburb_division(df_by_year, year)
  sum_by_suburb_year <- aggregate(Incidents.Recorded ~ Suburb.Town.Name + Offence.Division, df_by_year, sum)
  stacked_sum <- matrix(sum_by_suburb_year$Incidents.Recorded, nrow = num_of_divisions, ncol = num_of_suburbs, byrow = TRUE)
  
  title <- paste("Total crime division by suburb in", year)
  colours <- c("green", "yellow", "blue", "pink", "red", "forestgreen")
  
  barplot(stacked_sum, names.arg = unique(sum_by_suburb_year$Suburb.Town.Name), col = colours, main = title, las = 2, cex.names = 0.5)
  legend("topleft", rev(as.vector(unique(sum_by_suburb_year$Offence.Division))), fill = rev(colours), cex = 0.5)
  
  print(sum_by_suburb_year)
}
```

Counts of crime subdivision by suburb for each year:
```{r}
for (year in rev(sort(unique(df$Year.ending.December)))) {
  df_by_year <- df[df$Year.ending.December == year, ]
  df_by_year <- zero_missing_suburb_subdivision(df_by_year, year)
  sum_by_suburb_year <- aggregate(Incidents.Recorded ~ Suburb.Town.Name + Offence.Subdivision, df_by_year, sum)
  stacked_sum <- matrix(sum_by_suburb_year$Incidents.Recorded, nrow = num_of_subdivisions, ncol = num_of_suburbs, byrow = TRUE)
  
  title <- paste("Total crime subdivision by suburb in", year)
  colours <- c("green", "yellow", "blue", "pink", "chocolate", "cyan",
               "red", "chocolate", "hotpink", "lightgreen", "khaki")
  
  barplot(stacked_sum, names.arg = unique(sum_by_suburb_year$Suburb.Town.Name), col = colours, main = title, las = 2, cex.names = 0.5)
  legend("topleft", rev(as.vector(unique(sum_by_suburb_year$Offence.Subdivision))), fill = rev(colours), cex = 0.5)
  
  print(sum_by_suburb_year)
}
```
