---
title: "Crime rate by type across selected Victorian suburbs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 1600)
```

On this page I want to find out crime rates across the burbs weighted by unit population.

The population data is downloaded from the [2016 Census website](https://datapacks.censusdata.abs.gov.au/datapacks/). Choose "2016 Census Datapacks", "General Community Profile", "State Suburbs", and click "Vic".

## Steps

Read all suburb files and combine them into one big data frame
```{r}
source("./code/load_crime_data.R")
```

The offence divisions are:
```{r}
offence_divisions
```

The offence subdivisions are:
```{r}
offence_subdivisions
```

The suburbs are:
```{r}
suburbs
```

The suburb codes has been downloaded from the [ABS](http://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1270.0.55.003July%202016?OpenDocument) website. Alternatively, suburb codes can be discovered manually using the [Quickstats](http://www.abs.gov.au/websitedbs/D3310114.nsf/Home/2016%20QuickStats) feature.
```{r}
source("./code/suburb_lookup_functions.R")
```

Now we read in the 2016 census data for regression analysis.
```{r}
all_people <- read.csv("./census2016/2016 Census GCP State Suburbs for VIC/2016Census_G01_VIC_SSC.csv")
people <- data.frame()

for (f in files) {
  people <- rbind(people, all_people[all_people$SSC_CODE_2016 == lookup_ssc_code(f), ])
}

people <- cbind(Suburb.Area = 0, people)
people <- cbind(Suburb.Density = 0, people)
people <- cbind(Suburb.Name = files, people)

for (f in files) {
  ssc <- lookup_ssc_code(f)
  area <- lookup_suburb_area(f)
  num_of_people <- people[people$SSC_CODE_2016 == ssc, "Tot_P_P"]
  people[people$SSC_CODE_2016 == ssc, "Suburb.Area"] <- area
  people[people$SSC_CODE_2016 == ssc, "Suburb.Density"] <- (num_of_people / area)
}

for (f in files) {
  row <- people[people$Suburb.Name == f, ]
  print(paste(sep = "", f, " has ", row$Tot_P_P, " people in ", row$Suburb.Area, " km2. Density = ", row$Suburb.Density, " people/km2."))
}
```

Aggregate crime by division, as before. Here, I'm only looking at 2016 crime data because this is the year we have census data.
```{r}
source("./code/crime_functions.R")

year <- 2016
df_by_year <- df[df$Year.ending.December == year, ]
df_by_year <- zero_missing_suburb_division(df_by_year, year)
sum_by_suburb_year <- aggregate(Incidents.Recorded ~ Suburb.Town.Name + Offence.Division, df_by_year, sum)
```

We look at each crime division to see how they correlate with population.

## Crimes against the person
```{r}
people <- cbind(Crime.Against.Person.Rate = 0, people)
crime_against_person <- sum_by_suburb_year[sum_by_suburb_year$Offence.Division == "A Crimes against the person", ]
for (f in files) {
  n <- gsub("_", " ", f)
  ssc <- lookup_ssc_code(f)
  count <- crime_against_person[crime_against_person$Suburb.Town.Name == n, "Incidents.Recorded"]
  num_of_people <- people[people$SSC_CODE_2016 == ssc, "Tot_P_P"]
  people[people$SSC_CODE_2016 == ssc, "Crime.Against.Person.Rate"] <- (count / num_of_people) * 100000
}
barplot(people$Crime.Against.Person.Rate, names.arg = people$Suburb.Name, las = 2, cex.names = 0.5, main = "Crime against the person per 100k people by suburb", col = c("red"))

fit <- lm(people$Crime.Against.Person.Rate ~ people$Tot_P_P)
plot(people$Crime.Against.Person.Rate ~ people$Tot_P_P)
abline(fit, col="red")
res <- resid(fit)
plot(res)
abline(0,0, col="red")
summary(fit)
```

It looks like we have 3 powerful outliers - Sunshine, Broady, and Dandy. Let's remove them and see what happens.

```{r}
no_outlier <- people
no_outlier <- no_outlier[no_outlier$Suburb.Name != "SUNSHINE", ]
no_outlier <- no_outlier[no_outlier$Suburb.Name != "BROADMEADOWS", ]
no_outlier <- no_outlier[no_outlier$Suburb.Name != "DANDENONG", ]
barplot(no_outlier$Crime.Against.Person.Rate, names.arg = no_outlier$Suburb.Name, las = 2, cex.names = 0.5, main = "Crime against the person per 100k people without outlier", col = c("red"))

fit <- lm(no_outlier$Crime.Against.Person.Rate ~ no_outlier$Tot_P_P)
plot(no_outlier$Crime.Against.Person.Rate ~ no_outlier$Tot_P_P)
abline(fit, col="red")
res <- resid(fit)
plot(res)
abline(0,0, col="red")
summary(fit)
```

What about population density then?
```{r}
fit <- lm(no_outlier$Crime.Against.Person.Rate ~ no_outlier$Suburb.Density)
plot(no_outlier$Crime.Against.Person.Rate ~ no_outlier$Suburb.Density)
abline(fit, col="blue")
res <- resid(fit)
plot(res)
abline(0,0, col="blue")
summary(fit)

```
The regression model tells us population size has nothing to do with crime rate. Every suburb has, on average, 925 crimes against the person per 100,000 people in 2016 irrespective of population size. In addition, there's a mild but insignificant negative correlation between crime rate and population density. In other words, high density living is mildly correlated with lower crimes.

## Property and deception offences
Next we look at property and deception offences, which include break-in, vehicle theft, and property damage.

```{r}
people <- cbind(Property.Deception.Rate = 0, people)
property_and_deception <- sum_by_suburb_year[sum_by_suburb_year$Offence.Division == "B Property and deception offences", ]
for (f in files) {
  n <- gsub("_", " ", f)
  ssc <- lookup_ssc_code(f)
  count <- property_and_deception[property_and_deception$Suburb.Town.Name == n, "Incidents.Recorded"]
  num_of_people <- people[people$SSC_CODE_2016 == ssc, "Tot_P_P"]
  people[people$SSC_CODE_2016 == ssc, "Property.Deception.Rate"] <- (count / num_of_people) * 100000
}
barplot(people$Property.Deception.Rate, names.arg = people$Suburb.Name, las = 2, cex.names = 0.5, main = "Property and deception crime per 100k people by suburb", col = c("red"))

fit <- lm(people$Property.Deception.Rate ~ people$Tot_P_P)
plot(people$Property.Deception.Rate ~ people$Tot_P_P)
abline(fit, col="red")
res <- resid(fit)
plot(res)
abline(0,0, col="red")
summary(fit)

fit <- lm(people$Property.Deception.Rate ~ people$Suburb.Density)
plot(people$Property.Deception.Rate ~ people$Suburb.Density)
abline(fit, col="blue")
res <- resid(fit)
plot(res)
abline(0,0, col="blue")
summary(fit)

```

Again, property and deception crimes have no correlation with either population size or density.

How about income and unemployment?

## Does crime rate correlate with income and unemployment?

Let's load income data.
```{r}
all_income <- read.csv("./census2016/2016 Census GCP State Suburbs for VIC/2016Census_G02_VIC_SSC.csv")

people <- cbind(Weekly.Household.Income = 0, people)

for (f in files) {
  ssc <- lookup_ssc_code(f)
  people[people$SSC_CODE_2016 == ssc, "Weekly.Household.Income"] <- all_income[all_income$SSC_CODE_2016 == ssc, "Median_tot_hhd_inc_weekly"]
}

for (f in files) {
  row <- people[people$Suburb.Name == f, ]
  print(paste(sep = "", "Median weekly total household income in ", f, " is $", row$Weekly.Household.Income))
}
```

Run regression of crime vs income
```{r}
fit <- lm(people$Crime.Against.Person.Rate ~ people$Weekly.Household.Income)
plot(people$Crime.Against.Person.Rate ~ people$Weekly.Household.Income)
abline(fit, col="red")
res <- resid(fit)
plot(res)
abline(0,0, col="red")
summary(fit)

fit <- lm(people$Property.Deception.Rate ~ people$Weekly.Household.Income)
plot(people$Property.Deception.Rate ~ people$Weekly.Household.Income)
abline(fit, col="red")
res <- resid(fit)
plot(res)
abline(0,0, col="red")
summary(fit)
```

It looks like people living in rich suburbs are too precious to fight one another, but higher income doesn't stop property crimes.

How rich / poor are our selected suburbs?
```{r}
vic_mean_income <- mean(all_income$Median_tot_hhd_inc_weekly)
print(paste(sep = "", "Average weekly household income in Victoria is $", vic_mean_income))

our_suburb_income <- people$Weekly.Household.Income
print(paste(sep = "", "Average weekly household income in our selected suburbs is $", mean(our_suburb_income)))
print(paste(sep = "", "Median weekly household income in our selected suburbs is $", median(our_suburb_income)))

t.test(our_suburb_income, mu=vic_mean_income)

boxplot(our_suburb_income)
abline(h=vic_mean_income, col="red")
```

How about unemployment?
```{r}
all_employment <- read.csv("./census2016/2016 Census GCP State Suburbs for VIC/2016Census_G40_VIC_SSC.csv")

people <- cbind(Unemployment = 0, people)

for (f in files) {
  ssc <- lookup_ssc_code(f)
  people[people$SSC_CODE_2016 == ssc, "Unemployment"] <- all_employment[all_employment$SSC_CODE_2016 == ssc, "Percent_Unem_loyment_P"]
}

for (f in files) {
  row <- people[people$Suburb.Name == f, ]
  print(paste(sep = "", "Unemployment rate in ", f, " is ", row$Unemployment, "%."))
}
```

Run regression of crime vs unemployment
```{r}
fit <- lm(people$Crime.Against.Person.Rate ~ people$Unemployment)
plot(people$Crime.Against.Person.Rate ~ people$Unemployment)
abline(fit, col="red")
res <- resid(fit)
plot(res)
abline(0,0, col="red")
summary(fit)

fit <- lm(people$Property.Deception.Rate ~ people$Unemployment)
plot(people$Property.Deception.Rate ~ people$Unemployment)
abline(fit, col="red")
res <- resid(fit)
plot(res)
abline(0,0, col="red")
summary(fit)
```

When you're unemployed, you have more time to fight. Except in Clayton. Are they unemployed students who are busy reading instead?
